---
title       : Active Physician number Prediction
subtitle    : inference against on corresponding city population
author      : Shuo Zhao
job         : Data Scientist
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r echo=FALSE, warning=FALSE}
library(ggplot2)

readData <- function() {
  dat <- read.csv(file = './APPENC02.txt' , sep = ',', header = FALSE)
  names(dat)[c(5, 8, 9, 16, 17)] <- c('population', 'physicians', 'beds', 'income', 'region')
  dat[, 'population'] <- dat[, 'population'] / 1000
  # remove outliers
  dat <- dat[-(1:2),]
  dat
}
```

## Shiny App Intro

In any regular cities around the world, we need physicians. But how many physicians do we need in a particular city? The number of active physicians is expected to be related to total population, number of hospital beds, and total personal income etc. Here, I built a shiny application to apply linear regression model based on city's population to infer the number of active physicians for the city.

The data set I am using provides selected county demographic information (CDI) for 440 of the most populous counties in the United States. Each line of the data set has an identification number with a county name and state abbreviation and provides information on 14 variables for a single county. The information generally pertains to the years 1990 and 1992. There are 17 variables in the original dataset, we only use population and physicians in the application.

---

## Deployed Application

* You can find the deployed application from here: 
 + https://garfiny.shinyapps.io/project_part1

* The source code of the App & this slide shared in github:
 + https://github.com/garfiny/Developing-Data-Products-Prj

---

## Feature 1: Exploratory Analysis
```{r echo=FALSE, warning=FALSE, fig.width= 12}
library(ggplot2)
library(gridExtra)

data = readData()
n = dim(data)[1]
p_mean <- mean(data$population)
fit <- lm(data = data, physicians~population)

boxplot <- ggplot(data = data, aes(x = factor(region),
                                   y = population,
                                   fill = region)) +
  xlab("Regions") + ylab("Population (In Thousand)") +
  geom_boxplot() + guides(fill = FALSE)

histogram <- ggplot(data = data, aes(x = population)) +
  geom_histogram(aes(y = ..density..), 
                 binwidth = 50, 
                 colour = "green", 
                 fill = "white") +
  geom_density(alpha = 0.2, fill = "white")
  
plot1 <- ggplot(data = data, aes(x = population, y = physicians)) +
  geom_point() + geom_smooth(method=lm) + 
  xlab("Population (In Thousand)") + ylab("Physicians") +
  ggtitle("Scatter Plot") + theme_bw()

plot2 <- ggplot(fit, aes(.fitted, .resid)) + geom_point() + 
  stat_smooth(method="loess") + 
  geom_hline(yintercept=0, col="red", linetype="dashed") +
  xlab("Fitted values") + ylab("Residuals") +
  ggtitle("Residual vs Fitted Plot") + theme_bw()
grid.arrange(boxplot, histogram, plot1, plot2, ncol=2, nrow=2)
```

---

## Feature 2 - Regression Analysis
```{r }
predictions <- ceiling(predict(fit, newdata = data.frame(population = c(1000, 2000, 3000))))
```

- Linear Regression Model: #Physicians =  `r round(fit$coefficients[1], 3)` + `r round(fit$coefficients[2], 3)` * #Population
- Every `r round(fit$coefficients[2], 3)` thousand people population increase will generate 1 more physician.
- Given populations `r c(1000, 2000, 3000)`, our predictions respectively are `r predictions`

```{r }
predictions <- predict(fit, data.frame(population = data$population))
mse <- sum((data$physicians - predictions)^2) / (n - 2)
yh <- predict(fit, newdata = data.frame(population = c(200, 2000)))
s_yh <- sqrt(mse * (1/n + (c(200, 2000) - p_mean)^2/sum((data$population - p_mean)^2)))
conf_interval1 <- yh[1] + c(-1, 1) * qt(0.995, n - 2) * s_yh[1]
conf_interval2 <- yh[2] + c(-1, 1) * qt(0.995, n - 2) * s_yh[2]
```

- 99% Confidence Interval for 200K population: `r conf_interval1`
- 99% Confidence Interval for 2000K population: `r conf_interval2`


